---
title: "Project 1"
author: "Noreen Mayat"
output: html_document
---

*Question*: How do romantic relationships, experiences and interactions impact happiness across gender groups? 

In this project, I explore how romantic relationships, experiences and interactions impact happiness across gender groups of male and female, through computing the average word frequency for a bag of "romance"-related words for male and female, LDA topic modeling, and word2vec. 

Make sure to install all packages below before running any code; I installed them by running: 
```{r}
# install.packages('package') for each package listed.
```

Set ur directory. I set mine running this in the console:
```{r}
# setwd('Desktop/GitHub/ads-fall2023-project1-nm3224/doc/')
```
You may check your directory by running getwd() in the console.

```{r load_libraries, warning=FALSE, include=FALSE, echo=FALSE}
library(word2vec)
library(uwot)
library(tidytext)
library(tidyverse)
library(dplyr)
library(ggplot2)
library(ggrepel)
library(tm)
library(wordcloud)
library(udpipe)

#for sentiment analysis  

library(syuzhet)
library(lubridate)
library(scales)
library(reshape2)
```

Let's take a look at some happy moments, and the accompanying data:
```{r data, warning=FALSE, echo=FALSE}
pathfile <- '../output/processed_moments.csv'
cleaned_data <- read.csv(pathfile)
demo_data <- read.csv('../data/demographic.csv')
head(cleaned_data$cleaned_hm)
head(demo_data)
```

```{r process data, echo=FALSE}
# We need to first combine `demo_data$gender` to `cleaned_data`, join on `wid` to compare gender groups.
## Per gender 
gender_df <- merge(cleaned_data, demo_data)
```

Here is the frequency of all words in our documents: these are the top 10 with the highest frequencies.

```{r word frequencies, echo=FALSE}
# Then, we count the frequency of all words in our documents.
# for some reason this code doesn't work unless I execute the steps in this stackoverflow link: https://stackoverflow.com/questions/51295402/r-on-macos-error-vector-memory-exhausted-limit-reached

dtm <- TermDocumentMatrix(cleaned_data$text)
m <- as.matrix(dtm)
v <- sort(rowSums(m),decreasing=TRUE)
d <- data.frame(word = names(v),freq=v)
head(d, 10)
```

Now, let's compare the frequency of words, stratified by each gender group.
Note that we have roughly the same number of observations for each group, with slightly more testimonies from males. There are some 42,019 observations labeled female, and 57,597 observations labeled female. 
```{r male/female dataframes, echo=FALSE}
#FEMALE vs MALE
f_1 <- subset(gender_df, gender_df$gender == 'f')
m_2 <- subset(gender_df, gender_df$gender == 'm')
```

Here are the most frequent terms for females discussing happy moments:
```{r most frequent terms for females, echo=FALSE}

dtm_f <- TermDocumentMatrix(f_1$text)
m_f <- as.matrix(dtm_f)
v_f <- sort(rowSums(m_f),decreasing=TRUE)
d_f <- data.frame(word = names(v_f),freq=v_f)
head(d_f, 10)
```
Here are the most frequent terms for males discussing happy moments:
```{r most frequent terms for males, echo=FALSE}
#Most frequent terms for males discussing happy moments:

dtm_m <- TermDocumentMatrix(m_2$text)
m_m <- as.matrix(dtm_m)
v_m <- sort(rowSums(m_m),decreasing=TRUE)
d_m <- data.frame(word = names(v_m),freq=v_m)
head(d_m, 10)
```

I don't see any major discrepancies off the bat; it may be more useful to now zone in on a specific bag of words I came up with related to romance, relationships, intimacy, and partnership. 
```{r bag of words}
love <- c('wife', 'husband', 'kiss', 'date', 'boyfriend', 'girlfriend', 'fiance', 'fiancee', 'engaged', 'sex', 'sexual', 'dating', 'romance', 'romantic', 'spouse', 'partner', 'lover', 'marriage')
```
 
```{r overall frequency of bag of words, echo=FALSE}
#Overall frequency of romance across both categories
romance <- subset(d, word %in% love)

#Subsetting by gender
romance_male <- subset(d_m, word %in% love)
romance_male$Gender <- 'Male'

romance_female <- subset(d_f, word %in% love)
romance_female$Gender <- 'Female'

#Compute average frequency for word for each gender category by dividing frequency for gender category by overall frequency over all documents
romance_male$avg_freq <- romance_male$freq / romance$freq
romance_female$avg_freq <- romance_female$freq / romance$freq
```

Let's take a look at the average word frequencies (frequency for gender m vs. f / overall frequency) for romance related words for both genders:
```{r plot word frequencies for genders, echo=FALSE}
romance_df <- rbind(romance_male, romance_female) 

ggplot(romance_df, aes(fill=Gender, y=avg_freq, x=word)) + 
  geom_bar(position="dodge", stat="identity") +
  theme(axis.text.x = element_text(angle = 90))

ggsave(filename = file.path("../figs", 'romantic_words_plot.png'))
```

It seems men that the average frequency for romance-related words is higher in the male category than the female category. This means men discuss their romantic lives and relationships when talking about their happy moments more than women. This could imply that romantic relationships, interactions, and experiences are more impactful and related to male happiness than female happiness. 

## Leveraging word2vec for Partner-related words and happiness
Now, using word2vec: first, we convert the data to a list of characters to input into our model. 
```{r conversion, echo=FALSE}
## Overall vocabulary between happiness and animals
x <- tolower(cleaned_data$cleaned_hm)
cat(x[1])
```

Lemmatizing our text and using speech tag (verb, adverb, noun, adjective) will make representation easier (let's say we want to see all adjectives and nouns relative to the topic of animals).
Note this cell takes a while to run; followed the same lemmatization process as the TA.  
```{r speech tagging and lemmatizing, echo=FALSE, include=FALSE}
process_data <- function(x, n_topics){
  anno <- udpipe(x, "english", trace = 10, parallel.cores = 1)
  anno <- subset(anno, !is.na(lemma) & nchar(lemma) > 1 & !upos %in% "PUNCT")
  anno$text <- sprintf("%s//%s", anno$lemma, anno$upos)
  x <- paste.data.frame(anno, term = "text", group = "doc_id", collapse = " ")
  model <- word2vec(x = x$text, dim = n_topics, iter = 20, split = c(" ", ".\n?!"))
  embedding <- as.matrix(model)

  viz <- umap(embedding, n_neighbors = 15, n_threads = 2)
  rownames(viz) <- rownames(embedding)

  df <- data.frame(word = gsub("//.+", "", rownames(viz)),
  upos = gsub(".+//", "", rownames(viz)),
  x = viz[, 1], y = viz[, 2],
  stringsAsFactors = FALSE)
  df <- subset(df, upos %in% c("ADJ", "NOUN"))
  return(list('data'=df,'model'=model))
}

l <- process_data(x, 15)
```

We now want to get the words most similar to partner in the embedding and we compare them to the words most similar to happy to observe a relationship.
```{r most similar to partner, echo=FALSE}
lookslike2 <- predict(l$model, c("partner//NOUN"), type = "nearest", top_n = 10)
word_list_partner <- lookslike2$`partner//NOUN`$term2
word_list_happiness <- predict(l$model, c("happy//ADJ"), type = "nearest", top_n = 50)$`happy//ADJ`$term2
```

```{r slicing, echo=FALSE}
partner_df <- subset(l$data, rownames(l$data) %in% word_list_partner)
happiness_df <- subset(l$data, rownames(l$data) %in% word_list_happiness)
```

Let's analyze partner and happy associated words for our corpus.
## Overall 
```{r umap plot, echo=FALSE}
options(ggrepel.max.overlaps = Inf) 
ggplot(partner_df, aes(x = x, y = y, label = word, color='blue')) +
geom_text_repel() + theme_void() + geom_text_repel(data=happiness_df, aes(x=x, y=y, label=word, color='red'))
labs(title = "100 most similar words to partner with word2vec - umap")
ggsave(filename = file.path("../figs", 'word2vec_overall.png'))
```

## Per gender 
Now, let's look at each per gender. 
Note this cell also takes a while to run; followed the same process as the TA. Apologies!
```{r per gender, echo=FALSE, include=FALSE}
x_1 <- tolower(f_1$cleaned_hm)
x_2 <- tolower(m_2$cleaned_hm)

l_1 <- process_data(x_1, 15)
l_2 <- process_data(x_2, 15)

word_list_partner1 <- predict(l_1$model, c("partner//NOUN"), type = "nearest", top_n = 10)$`partner//NOUN`$term2
word_list_happiness1 <- predict(l_1$model, c("happy//ADJ"), type = "nearest", top_n = 50)$`happy//ADJ`$term2
word_list_partner2 <- predict(l_2$model, c("partner//NOUN"), type = "nearest", top_n = 10)$`partner//NOUN`$term2
word_list_happiness2 <- predict(l_2$model, c("happy//ADJ"), type = "nearest", top_n = 50)$`happy//ADJ`$term2

partner1 <- subset(l_1$data, rownames(l_1$data) %in% word_list_partner1)
happiness1 <- subset(l_1$data, rownames(l_1$data) %in% word_list_happiness1)
partner2 <- subset(l_2$data, rownames(l_2$data) %in% word_list_partner2)
happiness2 <- subset(l_2$data, rownames(l_2$data) %in% word_list_happiness2)
```

```{r umap plot with genres, warning=FALSE, echo=FALSE}
options(ggrepel.max.overlaps = Inf) 
ggplot(partner1, aes(x = x, y = y, label = word, color='blue')) +
geom_text_repel() + theme_void() + geom_text_repel(data=happiness1, aes(x=x, y=y, label=word, color='red')) +
geom_text_repel() + theme_void() + geom_text_repel(data=happiness2, aes(x=x, y=y, label=word, color='magenta')) +
geom_text_repel() + theme_void() + geom_text_repel(data=partner2, aes(x=x, y=y, label=word, color='cyan')) + scale_color_manual(values=c("blue", "red", "magenta", 'cyan'),
                          labels = c("women - partner", "men - partner", "men - happiness", 'women - happiness'))
labs(title = "Most similar words to partner with word2vec - umap")
ggsave('../figs/word2vec_genres.png')
```
From this analysis we can see that male happiness is very closely associated with partnership/partner words, while female happiness is not. 

## Sentiment Analysis
Now, let's conduct a sentiment analysis on text documents referring to "love" and see how the sentiments vary between men and women.
```{r sentiment analysis, echo=FALSE}
partner_text <- gender_df[grep(paste(love, collapse='|'), gender_df$text, ignore.case=TRUE),]
partner_mom <- iconv(partner_text$text)
s <- get_nrc_sentiment(partner_mom)
head(s)
```
We have 10 sentiments to look at.
Let's look at these sentiments for documents including the words listed in the "love" listed I provided.
### Overall
```{r plot figure, echo=FALSE}
barplot(colSums(s),
        las = 2,
        col = rainbow(10),
        ylab = 'Count',
        main = 'Sentiment Scores for Partner-Related Tweets')
```

```{r save image, echo=FALSE}
png("../figs/barplot_all.png")
dev.off()
```

Overall, sentiment related to partner text is positive and with joy. Let's test this analysis and compare for both genders.
## Per Gender
```{r gender sentiment analysis, echo=FALSE}
partner_text_male <- m_2[grep(paste(love, collapse='|'), m_2$text, ignore.case=TRUE),]
partner_male <- iconv(partner_text_male$text)
s_male <- get_nrc_sentiment(partner_male)

partner_text_female <- f_1[grep(paste(love, collapse='|'), f_1$text, ignore.case=TRUE),]
partner_female <- iconv(partner_text_female$text)
s_female <- get_nrc_sentiment(partner_female)
```

```{r sentiment analysis dataframe, echo=FALSE}
# Let's put all the data into one dataframe.
sent <- colSums(s_male)
df_m <- data.frame(sent, Gender = c('Male'))
df_m <- cbind(feeling = rownames(df_m), df_m)

sent <- colSums(s_female)
df_f <- data.frame(sent, Gender = c('Female'))
df_f <- cbind(feeling = rownames(df_f), df_f)

sents <- rbind(df_m, df_f)
```

Let's plot both side by side:
```{r plot sentiment analysis by gender, echo=FALSE}
ggplot(sents, aes(fill=Gender, y=sent, x=feeling)) + 
  geom_bar(position="dodge", stat="identity") +
  theme(axis.text.x = element_text(angle = 90))

ggsave(filename = file.path("../figs", 'sentiment_plot_genders.png'))
```
Overall, documents referring to what I defined as "partner" or "relationship" words have higher sentiment values in categories positive, joy, trust, and more for males than for females. 
