---
title: "Project 1"
author: "Noreen Mayat"
output: html_document
---

*Question*: How do romantic relationships, experiences and interactions impact happiness across gender groups? 

In this project, I explore how romantic relationships, experiences and interactions impact happiness across gender groups of male and female, through computing the average word frequency for a bag of "romance"-related words for male and female, LDA topic modeling, and word2vec. 

Make sure to install all packages below before running any code; I installed them by running: 
```{r}
#install.packages('package') for each package listed.
```

```{r load_libraries, warning=FALSE, echo=FALSE}
library(tidytext)
library(tidyverse)
library(dplyr)
library(ggplot2)
library(ggrepel)
library(plyr)
library(tm)
library(wordcloud)
```

```{r data, warning=FALSE, echo=FALSE}
pathfile <- '../output/processed_moments.csv'
cleaned_data <- read.csv(pathfile)
demo_data <- read.csv('https://raw.githubusercontent.com/megagonlabs/HappyDB/master/happydb/data/demographic.csv')
head(cleaned_data$cleaned_hm)
head(demo_data)
```


```{r}
#make all text lowercase
cleaned_data <- cleaned_data %>% 
  mutate(cleaned_hm = tolower(cleaned_hm))
```

We need to combine `demo_data$gender` to `cleaned_data`, join on `wid` to compare gender groups.

```{r process data, echo=FALSE}
## Per gender 
gender_df <- merge(cleaned_data, demo_data)
```

Let's count the frequency of all words in our documents first.

```{r}
#for some reason this code doesn't work unless I execute the steps in this stackoverflow link: https://stackoverflow.com/questions/51295402/r-on-macos-error-vector-memory-exhausted-limit-reached

dtm <- TermDocumentMatrix(cleaned_data$text)
m <- as.matrix(dtm)
v <- sort(rowSums(m),decreasing=TRUE)
d <- data.frame(word = names(v),freq=v)
head(d, 10)
```

Now, let's compare the frequency of words, stratified by each gender group.
Note that we have roughly the same number of observations for each group, with slightly more testimonies from males. There are some 42,019 observations labeled female, and 57,597 observations labeled female. 
```{r}
#FEMALE vs MALE
f_1 <- subset(gender_df, gender_df$gender == 'f')
m_2 <- subset(gender_df, gender_df$gender == 'm')
```

```{r}
#Most frequent terms for females discussing happy moments:

dtm_f <- TermDocumentMatrix(f_1$text)
m_f <- as.matrix(dtm_f)
v_f <- sort(rowSums(m_f),decreasing=TRUE)
d_f <- data.frame(word = names(v_f),freq=v_f)
head(d_f, 10)
```

```{r}
#Most frequent terms for males discussing happy moments:

dtm_m <- TermDocumentMatrix(m_2$text)
m_m <- as.matrix(dtm_m)
v_m <- sort(rowSums(m_m),decreasing=TRUE)
d_m <- data.frame(word = names(v_m),freq=v_m)
head(d_m, 10)
```

I don't see any major discrepancies off the bat; it may be more useful to now zone in on a specific bag of words.
```{r}
love <- c('wife', 'husband', 'kiss', 'date', 'boyfriend', 'girlfriend', 'fiance', 'fiancee', 'engaged', 'sex', 'sexual', 'dating', 'romance', 'romantic', 'spouse', 'partner', 'lover')
```
 
```{r}
#Overall frequency of romance across both categories
romance <- subset(d, word %in% love)

#Subsetting by gender
romance_male <- subset(d_m, word %in% love)
romance_male$Gender <- 'Male'

romance_female <- subset(d_f, word %in% love)
romance_female$Gender <- 'Female'

#Compute average frequency for word for each gender category by dividing frequency for gender category by overall frequency over all documents
romance_male$avg_freq <- romance_male$freq / romance$freq

romance_female$avg_freq <- romance_female$freq / romance$freq
```

```{r}
romance_df <- rbind(romance_male, romance_female) 

ggplot(romance_df, aes(fill=Gender, y=avg_freq, x=word)) + 
    geom_bar(position="dodge", stat="identity")
```

It seems men that the average frequency for romance-related words is higher in the male category than the female category. This means men discuss their romantic lives and relationships when talking about their happy moments more than women. This could imply that romantic relationships, interactions, and experiences are more impactful and related to male happiness than female happiness. 

